---
title: "POP247 - campus EDA"
subtitle: "Models for St Lucia - sites & modes of transport"
# author: "Radoslaw Panczak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
mainfont: DejaVu Sans
output: 
  html_document: 
    # css: custom.css
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: true
    theme: united 
    highlight: pygments 
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../docs") })
---

<!-- ------------------------------------------------------------ --> 
<!-- ------------------------------------------------------------ --> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      fig.width=8, fig.height=6, dpi=300, 
                      out.width="800px", out.height="600px",
                      cache=TRUE)

# knitr::opts_knit$set(root.dir = 'R:/POP247-Q0786/')
# setwd('R:/POP247-Q0786/')

set.seed(12345)
options(scipen = 999)

library(pacman) 
p_load(tidyverse, magrittr, lubridate,
       scales, gghighlight, fitdistrplus, MASS, mgcv, gratia, INLA, INLAutils)

inla.setOption(num.threads = 3) 
```

# People counts

```{r echo=FALSE}
campus_site_mode <- readRDS("data/counters/clean/uq_campus_long_mode.Rds") %>% 
  filter(  date >= ymd_hms("2016-08-15 00:00:00", tz = "Australia/Brisbane") & 
             date <= ymd_hms("2016-08-21 23:59:59", tz = "Australia/Brisbane"))

# table(campus_site_mode$direction)
# table(campus_site_mode$site, campus_site_mode$type)
# table(as.Date(campus_site_mode$date, tz = "Australia/Brisbane"))

campus_inbound <- bind_cols(date = rep(seq(ymd_hms("2016-08-15 00:00:00", 
                                                   tz = "Australia/Brisbane"), 
                                           ymd_hms("2016-08-21 23:59:00", 
                                                   tz = "Australia/Brisbane"), 
                                           as.difftime(15, units = "mins")), 9),
                            site = c(rep("Chancellors", 672), rep("ferry", 672),
                                     rep("green_bridge", 672), rep("green_bridge", 672),
                                     rep("lakes", 672), rep("Macquarie", 672),
                                     rep("McGregor", 672), rep("Schonell", 672), 
                                     rep("uq", 672)),
                            type = c(c(rep("vehicles", 672), rep("public_transport", 672),
                                       rep("cyclists", 672), rep("pedestrians", 672),
                                       rep("public_transport", 672), rep("cyclists", 672),
                                       rep("vehicles", 672), rep("vehicles", 672), 
                                       rep("public_transport", 672)))
) %>% 
  left_join(filter(campus_site_mode, direction == "inbound")) %>% 
  dplyr::select(-direction) %>%
  mutate(direction = "inbound") %>% 
  mutate(count = ifelse(is.na(count), 0, count)) %>% 
  mutate(time = as.numeric(date - min(campus_site_mode$date)) / 900) %>% 
  mutate(hod = as.numeric(strftime(date, "%H", tz = ""))) %>% 
  mutate(dow = as.numeric(strftime(date, "%d", tz = "")) - 14) %>% 
  mutate(wknd = ifelse(test = dow >=6 & dow <= 7, yes = 1, no = 0)) %>% 
  arrange(site, date)

# table(campus_site_mode$site, campus_site_mode$type)
# table(campus_inbound$site, campus_inbound$type)
# 
# table(as.Date(campus_site_mode$date, tz = "Australia/Brisbane"))
# table(as.Date(campus_inbound$date, tz = "Australia/Brisbane"))

campus_outbound <- bind_cols(date = rep(seq(ymd_hms("2016-08-15 00:00:00", 
                                                    tz = "Australia/Brisbane"), 
                                            ymd_hms("2016-08-21 23:59:00", 
                                                    tz = "Australia/Brisbane"), 
                                            as.difftime(15, units = "mins")), 9),
                             site = c(rep("Chancellors", 672), rep("ferry", 672),
                                      rep("green_bridge", 672), rep("green_bridge", 672),
                                      rep("lakes", 672), rep("Macquarie", 672),
                                      rep("McGregor", 672), rep("Schonell", 672), 
                                      rep("uq", 672)),
                             type = c(c(rep("vehicles", 672), rep("public_transport", 672),
                                        rep("cyclists", 672), rep("pedestrians", 672),
                                        rep("public_transport", 672), rep("cyclists", 672),
                                        rep("vehicles", 672), rep("vehicles", 672), 
                                        rep("public_transport", 672)))
) %>% 
  left_join(filter(campus_site_mode, direction == "outbound")) %>% 
  dplyr::select(-direction) %>%
  mutate(direction = "outbound") %>% 
  mutate(count = ifelse(is.na(count), 0, count)) %>% 
  mutate(time = as.numeric(date - min(campus_site_mode$date)) / 900) %>% 
  mutate(hod = as.numeric(strftime(date, "%H", tz = ""))) %>% 
  mutate(dow = as.numeric(strftime(date, "%d", tz = "")) - 14) %>% 
  mutate(wknd = ifelse(test = dow >=6 & dow <= 7, yes = 1, no = 0)) %>% 
  arrange(site, date)

campus <- bind_rows(campus_inbound, campus_outbound) 

# as character for better graphics
campus$site_type <- as.character(interaction(campus$type, campus$site, drop = TRUE, sep = " "))

rm(campus_site_mode)
```

```{r eval=FALSE, include=FALSE}
campus_wide <- full_join(campus_inbound  %>% rename(inbound = count),
                         campus_outbound %>% dplyr::select(date, site, type, count) %>% rename(outbound = count)) %>% 
  mutate(tally = inbound - outbound)
```

## Time series 

### All 

People counts, one full week, 15min intervals

```{r echo=FALSE}
campus %>% 
  mutate(count = ifelse(test = direction == "outbound", yes = -1 * count, no = count)) %>%
  ggplot(aes(x = date, y = count)) + 
  geom_col(aes(fill = direction)) + 
  scale_x_datetime() + theme_minimal() + 
  xlab("") + ylab("People* counts") +
  facet_wrap(~site_type)
```

```{r echo=FALSE}
campus %>% 
  mutate(count = ifelse(test = direction == "outbound", yes = -1 * count, no = count)) %>%
  ggplot(aes(x = date, y = count)) + 
  geom_line(aes(colour = direction)) + 
  scale_x_datetime() + theme_minimal() + 
  xlab("") + ylab("People* counts") +
  facet_wrap(~site_type)
```

### Cars

```{r echo=FALSE}
campus %>% 
  filter(site %in% c("Chancellors", "McGregor", "Schonell")) %>% 
  mutate(count = ifelse(test = direction == "outbound", yes = -1 * count, no = count)) %>%
  ggplot(aes(x = date, y = count)) + 
  geom_line(aes(color = site, group = interaction(site, direction))) + 
  scale_x_datetime() + theme_minimal() + 
  xlab("") + ylab("People* counts") # + geom_hline(yintercept = 0) 
```

Zero counts on Chancellors from 2016-08-18 10:15:00 to 2016-08-18 12:00:00

### Public transport

```{r echo=FALSE}
campus %>% 
  filter(site %in% c("ferry", "lakes", "uq")) %>% 
  mutate(count = ifelse(test = direction == "outbound", yes = -1 * count, no = count)) %>%
  ggplot(aes(x = date, y = count)) + 
  geom_line(aes(color = site, group = interaction(site, direction))) + 
  scale_x_datetime() + theme_minimal() + 
  xlab("") + ylab("People* counts") # + geom_hline(yintercept = 0) 
```


### Pedestrians & cyclists

```{r echo=FALSE}
campus %>% 
  filter(site %in% c("green_bridge", "Macquarie")) %>% 
  mutate(count = ifelse(test = direction == "outbound", yes = -1 * count, no = count)) %>%
  ggplot(aes(x = date, y = count)) + 
  geom_line(aes(color = site, group = interaction(site, direction))) + 
  scale_x_datetime() + theme_minimal() + 
  xlab("") + ylab("People* counts") +
  facet_wrap(~type)
```

green_bridge, pedestrians, outbound massive outliers:

2016-08-20 07:30:00
137

2016-08-20 07:45:00
265


## Distribution of counts

Frequency of counts from one week

```{r echo=FALSE}
campus %>% 
  ggplot(aes(x = count)) + 
  geom_histogram(aes(fill = direction), binwidth = 10) + 
  gghighlight() +
  theme_minimal() + 
  facet_wrap(~direction) +
  xlab("") + ylab("People* counts") 
```

## Distribution fits for inbound

```{r}
campus_inbound_poi <- fitdist(campus_inbound$count, "pois")
campus_inbound_nb <- fitdist(campus_inbound$count, "nbinom")

par(mfrow = c(1,2))
denscomp(list(campus_inbound_poi, campus_inbound_nb), legendtext = c("Poisson", "negative binomial"), fitlty = 1)
cdfcomp(list(campus_inbound_poi, campus_inbound_nb), legendtext = c("Poisson", "negative binomial"), fitlty = 1)

# gofstat(list(campus_inbound_poi, campus_inbound_nb), fitnames = c("Poisson", "negative binomial"))
```

```{r message=FALSE, warning=FALSE, include=FALSE}
dev.off()
rm(campus_inbound_poi, campus_inbound_nb)
```


# INLA

## Setup

```{r}
# ####
# Jeffreys prior 
a1 <- 5e-5
b1 <- 5e-5
lgprior1 <- list(prec = list(param = c(a1, b1)))

# Gelman prior
a2 <- -0.5
b2 <- 5e-5
lgprior2 <- list(prec = list(param = c(a2, b2)))

# iid prior 
# Schrödle & Held 2010 & Blangiardo et al 2013
a0 <- 1
b0 <- 0.1
prior.nu <- list(prec = list(param = c(a0, b0)))

# intercept & fixed
inla.set.control.fixed.default() 

# intercept ~ N(0,0) 
# other fixed effects ~ N(0, 0.001) 
# 
# where the format is N(mean, precision) 
# precision = inverse of the variance. 

# PC prior
U <- 1
hyper.prec = list(theta = list(
  prior = "pc.prec",
  param = c(U, 0.01)
))

# back to factor for regs
campus_inbound$site_type <- as.character(interaction(campus_inbound$type, campus_inbound$site, drop = TRUE, sep = " "))
campus_outbound$site_type <- as.character(interaction(campus_outbound$type, campus_outbound$site, drop = TRUE, sep = " "))

levels(campus_inbound$site_type)

```


## Inbound 

### RW1

#### One trend for all time series (rw1)

```{r}
mod1 <- inla(count ~ site_type + 
               f(time, model = "rw1", scale.model = TRUE, hyper = hyper.prec), 
             family = "nbinomial", 
             control.compute = list(dic= TRUE, waic = TRUE),  
             data = campus_inbound)
```

```{r}
summary(mod1)
autoplot(mod1)

P1 <- bind_cols(time = campus_inbound$time, count = campus_inbound$count, site_type = campus_inbound$site_type,
                fit = mod1$summary.fitted.values$`0.5quant`,
                lci = mod1$summary.fitted.values$`0.025quant`, 
                uci = mod1$summary.fitted.values$`0.975quant`)

ggplot(P1, aes(x = time)) +
  geom_line(aes(y = count), colour = "orange", alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.3) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

P1 %>% 
  filter(site_type == "public_transport lakes") %>% 
  ggplot(aes(x = time)) +
  geom_point(aes(y = count), colour = "orange", size = 2, alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

ggplot(P1, aes(x = count, y = fit)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("Counts")  + ylab("fitted")  
```

#### One trend for mode of transport (rw1)

```{r}
inla.setOption(scale.model.default = TRUE)

campus_inbound$cyclists <- as.numeric(campus_inbound$type == "cyclists")
campus_inbound$pedestrians <- as.numeric(campus_inbound$type == "pedestrians")
campus_inbound$public_transport <- as.numeric(campus_inbound$type == "public_transport")
campus_inbound$vehicles <- as.numeric(campus_inbound$type == "vehicles")

campus_inbound$time2 <- campus_inbound$time
campus_inbound$time3 <- campus_inbound$time
campus_inbound$time4 <- campus_inbound$time
  
mod2 <- inla(count ~ site_type + 
               f(time, cyclists, model = "rw1", hyper = hyper.prec) + 
               f(time2, pedestrians, model = "rw1", hyper = hyper.prec) + 
               f(time3, public_transport, model = "rw1", hyper = hyper.prec) + 
               f(time4, vehicles, model = "rw1", hyper = hyper.prec), 
             family = "nbinomial", 
             control.compute = list(dic= TRUE, waic = TRUE),  
             data = campus_inbound)
```

```{r}
summary(mod2)
autoplot(mod2)

P2 <- bind_cols(time = campus_inbound$time, count = campus_inbound$count, site_type = campus_inbound$site_type,
                fit = mod2$summary.fitted.values$`0.5quant`,
                lci = mod2$summary.fitted.values$`0.025quant`, 
                uci = mod2$summary.fitted.values$`0.975quant`)

ggplot(P2, aes(x = time)) +
  geom_line(aes(y = count), colour = "orange", alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.3) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

P2 %>% 
  filter(site_type == "public_transport lakes") %>% 
  ggplot(aes(x = time)) +
  geom_point(aes(y = count), colour = "orange", size = 2, alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

ggplot(P2, aes(x = count, y = fit)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("Counts")  + ylab("fitted")
  
```

#### One trend for site (rw1)

```{r}
campus_inbound$Chancellors <- as.numeric(campus_inbound$site == "Chancellors")
campus_inbound$ferry <- as.numeric(campus_inbound$site == "ferry")
campus_inbound$green_bridge <- as.numeric(campus_inbound$site == "green_bridge")
campus_inbound$lakes <- as.numeric(campus_inbound$site == "lakes")
campus_inbound$Macquarie <- as.numeric(campus_inbound$site == "Macquarie")
campus_inbound$McGregor <- as.numeric(campus_inbound$site == "McGregor")
campus_inbound$Schonell <- as.numeric(campus_inbound$site == "Schonell")
campus_inbound$uq <- as.numeric(campus_inbound$site == "uq")

campus_inbound$time2 <- campus_inbound$time
campus_inbound$time3 <- campus_inbound$time
campus_inbound$time4 <- campus_inbound$time
campus_inbound$time5 <- campus_inbound$time
campus_inbound$time6 <- campus_inbound$time
campus_inbound$time7 <- campus_inbound$time
campus_inbound$time8 <- campus_inbound$time
  
mod3 <- inla(count ~ site_type + 
               f(time, Chancellors, model = "rw1", hyper = hyper.prec) + 
               f(time2, ferry, model = "rw1", hyper = hyper.prec) + 
               f(time3, green_bridge, model = "rw1", hyper = hyper.prec) + 
               f(time4, lakes, model = "rw1", hyper = hyper.prec) + 
               f(time5, Macquarie, model = "rw1", hyper = hyper.prec) + 
               f(time6, McGregor, model = "rw1", hyper = hyper.prec) + 
               f(time7, Schonell, model = "rw1", hyper = hyper.prec) + 
               f(time8, uq, model = "rw1", hyper = hyper.prec), 
             family = "nbinomial", 
             control.compute = list(dic= TRUE, waic = TRUE),  
             data = campus_inbound)
```

```{r}
summary(mod3)
autoplot(mod3)

P3 <- bind_cols(time = campus_inbound$time, count = campus_inbound$count, site_type = campus_inbound$site_type,
                fit = mod3$summary.fitted.values$`0.5quant`,
                lci = mod3$summary.fitted.values$`0.025quant`, 
                uci = mod3$summary.fitted.values$`0.975quant`)

ggplot(P3, aes(x = time)) +
  geom_line(aes(y = count), colour = "orange", alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.3) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

P3 %>% 
  filter(site_type == "public_transport lakes") %>% 
  ggplot(aes(x = time)) +
  geom_point(aes(y = count), colour = "orange", size = 2, alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

ggplot(P3, aes(x = count, y = fit)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("Counts")  + ylab("fitted")
  
```

#### One trend for mode & site (rw1)

```{r}
campus_inbound$cyclists_green_bridge <- as.numeric(campus_inbound$site_type == "cyclists green_bridge")
campus_inbound$cyclists_Macquarie <- as.numeric(campus_inbound$site_type == "cyclists Macquarie")
campus_inbound$pedestrians_green_bridge <- as.numeric(campus_inbound$site_type == "pedestrians green_bridge")
campus_inbound$public_transport_ferry <- as.numeric(campus_inbound$site_type == "public_transport ferry")
campus_inbound$public_transport_lakes <- as.numeric(campus_inbound$site_type == "public_transport lakes")
campus_inbound$public_transport_uq <- as.numeric(campus_inbound$site_type == "public_transport uq")
campus_inbound$vehicles_Chancellors <- as.numeric(campus_inbound$site_type == "vehicles Chancellors")
campus_inbound$vehicles_McGregor <- as.numeric(campus_inbound$site_type == "vehicles McGregor")
campus_inbound$vehicles_Schonell <- as.numeric(campus_inbound$site_type == "vehicles Schonell")

campus_inbound$time9 <- campus_inbound$time
  
mod4 <- inla(count ~ site_type + 
               f(time, cyclists_green_bridge, model = "rw1", hyper = hyper.prec) + 
               f(time2, cyclists_Macquarie, model = "rw1", hyper = hyper.prec) + 
               f(time3, pedestrians_green_bridge, model = "rw1", hyper = hyper.prec) + 
               f(time4, public_transport_ferry, model = "rw1", hyper = hyper.prec) + 
               f(time5, public_transport_lakes, model = "rw1", hyper = hyper.prec) + 
               f(time6, public_transport_uq, model = "rw1", hyper = hyper.prec) + 
               f(time7, vehicles_Chancellors, model = "rw1", hyper = hyper.prec) + 
               f(time8, vehicles_McGregor, model = "rw1", hyper = hyper.prec) + 
               f(time9, vehicles_Schonell, model = "rw1", hyper = hyper.prec), 
             family = "nbinomial", 
             control.compute = list(dic= TRUE, waic = TRUE),  
             data = campus_inbound)
```

```{r}
summary(mod4)
autoplot(mod4)

P4 <- bind_cols(time = campus_inbound$time, count = campus_inbound$count, site_type = campus_inbound$site_type,
                fit = mod4$summary.fitted.values$`0.5quant`,
                lci = mod4$summary.fitted.values$`0.025quant`, 
                uci = mod4$summary.fitted.values$`0.975quant`)

ggplot(P4, aes(x = time)) +
  geom_line(aes(y = count), colour = "orange", alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.3) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

P4 %>% 
  filter(site_type == "public_transport lakes") %>% 
  ggplot(aes(x = time)) +
  geom_point(aes(y = count), colour = "orange", size = 2, alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

ggplot(P4, aes(x = count, y = fit)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("Counts")  + ylab("fitted")
```

### RW2 

#### One trend for all time series (rw2)

```{r}
mod5 <- inla(count ~ site_type + 
               f(time, model = "rw2", scale.model = TRUE, hyper = hyper.prec), 
             family = "nbinomial", 
             control.compute = list(dic= TRUE, waic = TRUE),  
             data = campus_inbound)
```

```{r}
summary(mod5)
autoplot(mod5)

P5 <- bind_cols(time = campus_inbound$time, count = campus_inbound$count, site_type = campus_inbound$site_type,
                fit = mod5$summary.fitted.values$`0.5quant`,
                lci = mod5$summary.fitted.values$`0.025quant`, 
                uci = mod5$summary.fitted.values$`0.975quant`)

ggplot(P5, aes(x = time)) +
  geom_line(aes(y = count), colour = "orange", alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.3) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

P5 %>% 
  filter(site_type == "public_transport lakes") %>% 
  ggplot(aes(x = time)) +
  geom_point(aes(y = count), colour = "orange", size = 2, alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

ggplot(P5, aes(x = count, y = fit)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("Counts")  + ylab("fitted")  
```

#### One trend for mode of transport (rw2)

```{r}
inla.setOption(scale.model.default = TRUE)

campus_inbound$cyclists <- as.numeric(campus_inbound$type == "cyclists")
campus_inbound$pedestrians <- as.numeric(campus_inbound$type == "pedestrians")
campus_inbound$public_transport <- as.numeric(campus_inbound$type == "public_transport")
campus_inbound$vehicles <- as.numeric(campus_inbound$type == "vehicles")

campus_inbound$time2 <- campus_inbound$time
campus_inbound$time3 <- campus_inbound$time
campus_inbound$time4 <- campus_inbound$time
  
mod6 <- inla(count ~ site_type + 
               f(time, cyclists, model = "rw2", hyper = hyper.prec) + 
               f(time2, pedestrians, model = "rw2", hyper = hyper.prec) + 
               f(time3, public_transport, model = "rw2", hyper = hyper.prec) + 
               f(time4, vehicles, model = "rw2", hyper = hyper.prec), 
             family = "nbinomial", 
             control.compute = list(dic= TRUE, waic = TRUE),  
             data = campus_inbound)
```

```{r}
summary(mod6)
autoplot(mod6)

P6 <- bind_cols(time = campus_inbound$time, count = campus_inbound$count, site_type = campus_inbound$site_type,
                fit = mod6$summary.fitted.values$`0.5quant`,
                lci = mod6$summary.fitted.values$`0.025quant`, 
                uci = mod6$summary.fitted.values$`0.975quant`)

ggplot(P6, aes(x = time)) +
  geom_line(aes(y = count), colour = "orange", alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.3) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

P6 %>% 
  filter(site_type == "public_transport lakes") %>% 
  ggplot(aes(x = time)) +
  geom_point(aes(y = count), colour = "orange", size = 2, alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

ggplot(P6, aes(x = count, y = fit)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("Counts")  + ylab("fitted")
  
```

#### One trend for site (rw2)

```{r}
campus_inbound$Chancellors <- as.numeric(campus_inbound$site == "Chancellors")
campus_inbound$ferry <- as.numeric(campus_inbound$site == "ferry")
campus_inbound$green_bridge <- as.numeric(campus_inbound$site == "green_bridge")
campus_inbound$lakes <- as.numeric(campus_inbound$site == "lakes")
campus_inbound$Macquarie <- as.numeric(campus_inbound$site == "Macquarie")
campus_inbound$McGregor <- as.numeric(campus_inbound$site == "McGregor")
campus_inbound$Schonell <- as.numeric(campus_inbound$site == "Schonell")
campus_inbound$uq <- as.numeric(campus_inbound$site == "uq")

campus_inbound$time2 <- campus_inbound$time
campus_inbound$time3 <- campus_inbound$time
campus_inbound$time4 <- campus_inbound$time
campus_inbound$time5 <- campus_inbound$time
campus_inbound$time6 <- campus_inbound$time
campus_inbound$time7 <- campus_inbound$time
campus_inbound$time8 <- campus_inbound$time
  
mod7 <- inla(count ~ site_type + 
               f(time, Chancellors, model = "rw2", hyper = hyper.prec) + 
               f(time2, ferry, model = "rw2", hyper = hyper.prec) + 
               f(time3, green_bridge, model = "rw2", hyper = hyper.prec) + 
               f(time4, lakes, model = "rw2", hyper = hyper.prec) + 
               f(time5, Macquarie, model = "rw2", hyper = hyper.prec) + 
               f(time6, McGregor, model = "rw2", hyper = hyper.prec) + 
               f(time7, Schonell, model = "rw2", hyper = hyper.prec) + 
               f(time8, uq, model = "rw2", hyper = hyper.prec), 
             family = "nbinomial", 
             control.compute = list(dic= TRUE, waic = TRUE),  
             data = campus_inbound)
```

```{r}
summary(mod7)
autoplot(mod7)

P7 <- bind_cols(time = campus_inbound$time, count = campus_inbound$count, site_type = campus_inbound$site_type,
                fit = mod7$summary.fitted.values$`0.5quant`,
                lci = mod7$summary.fitted.values$`0.025quant`, 
                uci = mod7$summary.fitted.values$`0.975quant`)

ggplot(P7, aes(x = time)) +
  geom_line(aes(y = count), colour = "orange", alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.3) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

P7 %>% 
  filter(site_type == "public_transport lakes") %>% 
  ggplot(aes(x = time)) +
  geom_point(aes(y = count), colour = "orange", size = 2, alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

ggplot(P7, aes(x = count, y = fit)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("Counts")  + ylab("fitted")
  
```

#### One trend for mode & site (rw2)

```{r}
campus_inbound$cyclists_green_bridge <- as.numeric(campus_inbound$site_type == "cyclists green_bridge")
campus_inbound$cyclists_Macquarie <- as.numeric(campus_inbound$site_type == "cyclists Macquarie")
campus_inbound$pedestrians_green_bridge <- as.numeric(campus_inbound$site_type == "pedestrians green_bridge")
campus_inbound$public_transport_ferry <- as.numeric(campus_inbound$site_type == "public_transport ferry")
campus_inbound$public_transport_lakes <- as.numeric(campus_inbound$site_type == "public_transport lakes")
campus_inbound$public_transport_uq <- as.numeric(campus_inbound$site_type == "public_transport uq")
campus_inbound$vehicles_Chancellors <- as.numeric(campus_inbound$site_type == "vehicles Chancellors")
campus_inbound$vehicles_McGregor <- as.numeric(campus_inbound$site_type == "vehicles McGregor")
campus_inbound$vehicles_Schonell <- as.numeric(campus_inbound$site_type == "vehicles Schonell")

campus_inbound$time9 <- campus_inbound$time
  
mod8 <- inla(count ~ site_type + 
               f(time, cyclists_green_bridge, model = "rw2", hyper = hyper.prec) + 
               f(time2, cyclists_Macquarie, model = "rw2", hyper = hyper.prec) + 
               f(time3, pedestrians_green_bridge, model = "rw2", hyper = hyper.prec) + 
               f(time4, public_transport_ferry, model = "rw2", hyper = hyper.prec) + 
               f(time5, public_transport_lakes, model = "rw2", hyper = hyper.prec) + 
               f(time6, public_transport_uq, model = "rw2", hyper = hyper.prec) + 
               f(time7, vehicles_Chancellors, model = "rw2", hyper = hyper.prec) + 
               f(time8, vehicles_McGregor, model = "rw2", hyper = hyper.prec) + 
               f(time9, vehicles_Schonell, model = "rw2", hyper = hyper.prec), 
             family = "nbinomial", 
             control.compute = list(dic= TRUE, waic = TRUE),  
             data = campus_inbound)
```

```{r}
summary(mod8)
autoplot(mod8)

P8 <- bind_cols(time = campus_inbound$time, count = campus_inbound$count, site_type = campus_inbound$site_type,
                fit = mod8$summary.fitted.values$`0.5quant`,
                lci = mod8$summary.fitted.values$`0.025quant`, 
                uci = mod8$summary.fitted.values$`0.975quant`)

ggplot(P8, aes(x = time)) +
  geom_line(aes(y = count), colour = "orange", alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  geom_ribbon(aes(ymin = lci, ymax = uci), alpha = 0.3) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

P8 %>% 
  filter(site_type == "public_transport lakes") %>% 
  ggplot(aes(x = time)) +
  geom_point(aes(y = count), colour = "orange", size = 2, alpha = 0.5) +
  geom_line(aes(y = fit), col = "purple") +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("") 

ggplot(P8, aes(x = count, y = fit)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~site_type) +
  theme_minimal() + xlab("Counts")  + ylab("fitted")
```

### Comparing

```{r}
dotchart(c(mod1$dic$dic, mod2$dic$dic, mod3$dic$dic, mod4$dic$dic,
           mod5$dic$dic, mod6$dic$dic, mod7$dic$dic, mod8$dic$dic),
         labels = c("mod1", "mod2", "mod3", "mod4", "mod5", "mod6", "mod7", "mod8"), main = "DIC")

dotchart(c(mod1$waic$waic, mod2$waic$waic, mod3$waic$waic, mod4$waic$waic,
           mod5$waic$waic, mod6$waic$waic, mod7$waic$waic, mod8$waic$waic),
         labels = c("mod1", "mod2", "mod3", "mod4", "mod5", "mod6", "mod7", "mod8"), main = "WAIC")
```



